---
title: "eds222 final"
author: "Emily Cobb"
date: "2023-11-30"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Package loading and data setup

```{r packages}
# Load all the packages needed here
library(tidyverse)
library(readr)
library(modelsummary) #to make nice tables and dataskim
library(here) #for file path simplicity
library(broom.mixed)
library(effectsize)
library(corrplot) #correlation matrix
library(leaflet) #map time
library(geojsonio)
library(rempsyc)
library(htmlwidgets)
library(webshot)
library(mapview)
```

Data import and cleaning

```{r}
hurricane <- read_csv("hurricanedisplacement2005.csv")
```

```{r}
#fix data errors and change names to their abv
hurricane<- hurricane %>% 
mutate(across(where(is.numeric), ~abs(.))) %>%
  mutate(
   evac.state  = case_when(
      str_detect(evac.state, "GEORIGA") ~ "GA",
       TRUE ~ evac.state  # Keep the original value if none of the conditions are met
    )) %>%
  mutate(
   state.most.time  = case_when(
      str_detect(state.most.time, "OKLAHOMA") ~ "OK",
      str_detect(state.most.time, "FLA") ~ "FL",
       TRUE ~ state.most.time ))
```

```{r}
#change to factor variables 
  hurricane <- hurricane %>% mutate(across(where(is.numeric), as.factor))
```

```{r}
#make new outcome variable binary-- did leave state or did not 
hurricane <- hurricane %>% mutate("didmove"= ifelse(lives.NO.still == c("2","3"), 1, 0 ))
# labels for plots  
hurricane <- hurricane %>% mutate("didmove2"= ifelse(lives.NO.still == c("1"), "Moved Away", "Moved Back" ))
```

```{r}
#fix race and work variables 
hurricane <- hurricane %>% 
  mutate("white" = ifelse(white == c("1"), "1", "0" )) %>%
  mutate("black" = ifelse(black == c("2"), "1", "0" )) %>%
  mutate("hispanic" = ifelse(hispanic == c("3"), "1", "0" )) %>%
  mutate("partime" = ifelse(partime == c("2"), "1", "0" )) %>%
  mutate("selfempfull" = ifelse(selfempfull == c("3"), "1", "0" )) %>%
  mutate("laidoff" = ifelse(laidoff == c("5"), 1, 0 )) %>% 
  mutate("selfemppart" = ifelse(selfemppart == c("4"), 1, 0 )) %>% 
  mutate("unemployed" = ifelse(unemployed == c("6"), 1, 0 )) %>% 
  mutate("disabled" = ifelse(disabled== c("7"), 1, 0 )) %>% 
  mutate("student" = ifelse(student == c("9"), 1, 0))
  hurricane <- hurricane %>% mutate(across(where(is.numeric), as.character)) 
```

## Plotting

```{r}
#summary stats and overview
datasummary_balance(~ 1, data = hurricane, caption = "summary of dataset")
```

```{r}
#summary of did move vs moved groups 

hurricanestats <- hurricane %>% select(
  didmove2, 
  `Black` = black,
  "Hispanic" = hispanic, 
  "White" = white,
  "Level of School Completed" = schooling, 
  "Employed Full-Time" = fulltime, 
  "Employed Part-Time" = partime, 
  "Self Employed Full-Time" = selfempfull,
  "Self Employed Part-Time" = selfemppart,
  "Laid off from Job" = laidoff, 
  "Unemployed" = unemployed,
  "Disabled" = disabled, 
  "Student" = student, 
  "Own Home" = own.rent, 
  "Lived in Apartment" = apartment, 
  "Lived in Single Family House" = singlefam, 
  "Lived in a Duplex" = duplex,
  "Damage to Residence" = res.damage,
  "Income" = income
)

caption <- "Compairison of Sociodemogrpahic and Housing/Hurricane Displacement Variables Between Moved and Moved Back Groups"

datasummary_balance(~didmove2,
                    data = hurricanestats,
                    title = caption)
                
```

```{r}
#make summary of evacuation states for chloropelth map
evac.summary <- hurricane %>% 
  group_by(evac.state) %>%
  summarise( people = n())
```

```{r}
#change names in state variable to merge better
evac.summary <- evac.summary %>%
  mutate(
   evac.state  = case_when(
      str_detect(evac.state, "GA") ~ "Georgia",
      str_detect(evac.state, "AL") ~ "Alabama",
      str_detect(evac.state, "FL") ~ "Florida",
      str_detect(evac.state, "LA") ~ "Louisiana",
      str_detect(evac.state, "MA") ~ "Massachusetts",
      str_detect(evac.state, "MS") ~ "Mississippi",
      str_detect(evac.state, "NC") ~ "North Carolina",
      str_detect(evac.state, "TN") ~ "Tennessee",
      str_detect(evac.state, "TX") ~ "Texas",
      str_detect(evac.state, "AR") ~ "Arkansas",
       TRUE ~ evac.state  # Keep the original value if none of the conditions are met
    ))

```

map making

```{r , eval = FALSE}
#map
states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp")

merged_data <- merge(states, evac.summary, by.x = "name", by.y = "evac.state", x.all = TRUE)


# Create a Leaflet map
map <- leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%
  setView(lng = -96, lat = 37.8, zoom = 4)  # Adjust the view as needed

#map
map <- map %>%
  addPolygons(
    data = merged_data,
    fillColor = ~colorNumeric("YlOrRd", people)(people),
    fillOpacity = 0.7,
    color = "white",
    weight = 1,
    highlight = highlightOptions(
      weight = 2,
      color = "#666",
      fillOpacity = 0.7,
      bringToFront = TRUE
    ),
    label = ~paste(name, "<br>", "people: ", people, sep = "")
  )

# Add a legend
map <- map %>%
  addLegend(
    data = merged_data,
    position = "bottomright",
    pal = colorNumeric("YlOrRd", merged_data$people),
    values = ~people,
    title = "Number of People Evacuated to:",
    opacity = 1
  )

# Display the map
map
```

save map

```{r, eval = FALSE}
mapshot(map, file = "map.png")
```

## Models

### logistic regression

```{r}

```

```{r}
#plot to visualize outcome
gg <- ggplot(data = hurricane, aes(x = income, y = didmove)) + 
  geom_jitter(width = 0, height = 0.05, alpha = 0.8) +
   labs(x = "income", y = "Moved out of NO") 
gg
```

```{r}
#filtering and fixing for model setup

hurricane <- hurricane %>% mutate_at(c("age"), as.numeric) %>% mutate_at(c("white", "hispanic", "black", "fulltime", "partime", "selfempfull", "selfemppart", "laidoff", "unemployed", "disabled","student", "apartment", "singlefam", "duplex", "own.rent"), as.character) 

hurricane <- hurricane %>% filter(!is.na(didmove))

#for everything
hurricane <- hurricane %>% mutate(across((white:hispanic | fulltime:student | apartment:duplex),~replace_na(.x, "0"))) %>% mutate(didmove = as.factor(didmove))

```

Models (added step wise to address OBV)

```{r}
#adding variables by catagory 
#age test
hurrmodel1 <- glm(didmove ~ age, family = "binomial", data = hurricane)
#more demographics
hurrmodel2 <- glm(didmove ~ age + sex + white + black + hispanic, family = "binomial", data = hurricane)
# + job and income
hurrmodel3 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income, family = "binomial", data = hurricane)
# + housing factors
hurrmodel4 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before  + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income + apartment + singlefam + duplex + own.rent, family = "binomial", data = hurricane)
# + hurricane factors 
hurrmodel5 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income + apartment + singlefam + duplex + own.rent + evac.state + res.damage, family = "binomial", data = hurricane)
# add interaction effect to test
hurrmodel6 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income + apartment + singlefam + duplex + own.rent + evac.state + res.damage + own.rent*res.damage, family = "binomial", data = hurricane)

model_list = list(   
                "age only simple model" = hurrmodel1,
                "adding race and demogrpahic factors" = hurrmodel2,
                "adding job and income"= hurrmodel3,
                "with housing factors" = hurrmodel4,
                "hurricane factors"= hurrmodel5, 
                "with res dmg x own or rent interaction" = hurrmodel6)
modelsummary(model_list, stars = TRUE, caption = "Regression models of variables predicting move or did not move back to NO after Hurricane Katrina with Standard Errors")

```

```{r}
# runnign again with matched sample sizes 
hurricane2 <- augment(hurrmodel6) 
#age test
hurrmodel1 <- glm(didmove ~ age, family = "binomial", data = hurricane)
#more demographics
hurrmodel2 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before, family = "binomial", data = hurricane2)
# + job and income
hurrmodel3 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income, family = "binomial", data = hurricane2)
# + housing factors
hurrmodel4 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before  + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income + apartment + singlefam + duplex + own.rent, family = "binomial", data = hurricane2)
# + hurricane factors 
hurrmodel5 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income + apartment + singlefam + duplex + own.rent + evac.state + res.damage, family = "binomial", data = hurricane)
# add interaction effect to test
hurrmodel6 <- glm(didmove ~ age + sex + pre_marriagestat + white + black + hispanic + health.before + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + income + apartment + singlefam + duplex + own.rent + evac.state + res.damage + own.rent*res.damage, family = "binomial", data = hurricane)

model_list = list(   
                "age only simple model" = hurrmodel1,
                "adding race and demogrpahic factors" = hurrmodel2,
                "adding job and income"= hurrmodel3,
                "with housing factors" = hurrmodel4,
                "hurricane factors"= hurrmodel5, 
                "with res dmg x own or rent interaction" = hurrmodel6)
modelsummary(model_list, stars = TRUE, caption = "Regression models of variables predicting move or did not move back to NO after Hurricane Katrina")
```

model with just demographic and hurricane variables of interest + a few control variables

```{r}
#model
hurrmodel7 <- glm(didmove ~ age + sex  + white + black + hispanic + income + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + own.rent + evac.state + res.damage + own.rent*res.damage, family = "binomial", data = hurricane)
modelsummary(("Model with interaction effect" = hurrmodel7), stars = TRUE, caption = "risk factors, controls")
```

```{r}
#no interaction 
hurrmodel8 <- glm(didmove ~ age + sex  + white + black + hispanic + income + schooling + fulltime + partime + selfempfull + selfemppart + laidoff + unemployed + disabled + student + own.rent + evac.state + res.damage, family = "binomial", data = hurricane)
model_list = list(
  " With Interaction Between owning/Renting and Residance Damage" = hurrmodel7, 
   "Without Interaction"= hurrmodel8           )
modelsummary(model_list, stars = TRUE, caption = "Table 3. Multivariate Logistic Regression Model Predicting Moved Away/Moved Back to NO with Standard Errors", output = "logisitcregressiontable.png")
```

```{r}
#odds ratio making 
odds <-exp(coef(hurrmodel8))

exp(cbind(Odds_Ratio = coef(hurrmodel8), confint(hurrmodel8)))
```

```{r}
# for interaction model
exp(cbind(Odds_Ratio = coef(hurrmodel7), confint(hurrmodel7)))
```

Analysis 2

t-test for a difference in health from those who moved vs those who didn't move

q37 is a general stress measure for katrina

q31-36 depression and anxiety qs, lowest scores = highest depression

```{r}
# testing didmove and depression score
hurricane <- hurricane %>% filter(!is.na(didmove))

#calc depression score 
hurricane <- hurricane %>% mutate(across(q31:q37, ~as.numeric(.x))) 

x <- hurricane %>% select(q31:q36)

hurricane3 <- hurricane %>% 
  mutate(depressionscore = rowSums(x))

#health score

hurricane3 <- hurricane3 %>% mutate(across(health.before:health.after, ~as.numeric(.x)))

hurricane4 <- hurricane3 %>% 
  mutate(healthscore =(health.after - health.before))


```

t tests

comparison of means needs to be troubleshoot

```{r}

q37means <- hurricane3 %>% group_by(didmove2) %>%
  summarise( "mean HK stress score" = mean(q37, na.rm=TRUE))
knitr::kable(q37means, caption = "Mean Stress Scores by Group")

#new cuter table thats why cooler and i like it more i hope 
datasummary((`Stress Index` = q37) + (`Depression Index` = depressionscore) + (`Health Change` = healthscore) ~ didmove2 * (mean*Arguments(na.rm=TRUE)),
            data = hurricane4, caption = "Compairison of Means Between Groups")
```

```{r}
# general stress
t.test <- t.test( q37 ~ didmove, data = hurricane3, na.rm = TRUE)

t.stress <- nice_t_test(
  data = hurricane3, 
  response = "q37",
  group = "didmove",
  warning = FALSE
)
 stress.table <- nice_table(t.stress)
 stress.table
 t.test
```

depression t-tests

means table

```{r}
demeans <- hurricane3 %>% group_by(didmove2) %>%
  summarise( "mean depression score" = mean(depressionscore, na.rm=TRUE))
knitr::kable(demeans, caption = "Mean Depression Scores by Group")
```

```{r}
# general stress
t.test2 <- t.test( depressionscore ~ didmove, data = hurricane3, na.rm = TRUE)

t.depression <- nice_t_test(
  data = hurricane3, 
  response = "depressionscore",
  group = "didmove",
  warning = FALSE
)
 depression.table <- nice_table(t.depression)
 depression.table
 t.test2
```

```{r}

#test for health change variable 
t.test3 <- t.test( healthscore ~ didmove, data = hurricane4, na.rm = TRUE)

t.health <- nice_t_test(
  data = hurricane4, 
  response = "healthscore",
  group = "didmove",
  warning = FALSE
)
 health.table <- nice_table(t.health)
 health.table
 modelsummary(t.test3,  output = "table.tex")
```

final t test table

```{r}
new_col <- data.frame(`P Value` = c(.002, .012, .082), `95 Percent CIs` = c("[-1.02 - 0.16]","[0.05 - 0.91]","[-0.49 - 0.36]"))


#new cuter table thats why cooler and i like it more i hope 
datasummary((`Stress Index` = q37) + (`Depression Index` = depressionscore) + (`Health Change` = healthscore) ~ didmove2 * (mean*Arguments(na.rm=TRUE)),
            data = hurricane4, 
            caption = "T-tests Between Moved Away and Moved Back", 
            add_columns = (new_col ), 
            digits = 4, 
            output = "ttests.png")
```
